Drug sensitivity prediction with normal inverse Gaussian shrinkage informed by external data

### 2020-04-28 docs/manuscript.Rnw
changed README file

  To summarise, estimation of only $\bm{\alpha}_{\text{feat}}$ or
  $\bm{\alpha}_{\text{drug}}$ (and consequently $\phi_{jd}$ and $\chi_{d}$)
  by NIG$_{\text{f}}$ and NIG$_{\text{d}}$, respectively is relatively unbiased,
  as evident from simulation Scenarios 1 and 2 (Figures 
  \ref{fig:simulations_gdsc_est1} and \ref{fig:simulations_gdsc_est2}). 
  In contrast, simultaneous 
  estimation in Scenario 3 results in overestimated $\phi_{jd}$ and 
  underestimated $\chi_{d}$ by NIG$_{\text{f}+\text{d}}$
  (Figure \ref{fig:simulations_gdsc_est3}). We conjecture that this interplay
  of drug and feature variance components is due to near-unidentifiability.e
  r, the resultconsequences are limited, since theiances on the $\beta_{jd}$ level are unbiasleft ed 
  (Figure \ref{fig:simulations_gdsc_var}). If separately estimated, scale 
  parameters $\lambda_{\text{feat}}$ and $\lambda_{\text{drug}}$ are 
  overestimated by NIG$_{\text{f}}$ and NIG$_{\text{d}}$, respectively.
  If estimated simultaneously NIG$_{\text{f}+\text{d}}$ overestimates 
  $\lambda_{\text{feat}}$ and underestimated $\lambda_{\text{drug}}$. On the
  $\beta_{jd}$ level, $\bm{\lambda}$ influences kurtoses 
  $\mathcal{K}(\beta_{jd})$. Figure \ref{fig:simulations_gdsc_kurt} shows the
  mean estimated kurtoses versus their true values. Kurtoses seem to be
  underestimated by NIG$_{\text{f}+\text{d}}$. In contrast, 
  NIG$_{\text{f}+\text{d}}^{-1}$ overestimates the true mean of the
  $\mathcal{K}(\beta_{jd})$, \Sexpr{true.kurtb3}, at \Sexpr{kurtb3}.
  Lastly, the results from Scenario 3 in SM Section \ref{sm-sec:simulations} 
  show that the VB approximation is quite good as compared to standard MCMC.
<<simulations_gdsc_kurt, fig.cap="Simulation results for Scenario 3: mean estimated prior variances $\\V(\\beta_{jd})$ versus true values, with line of identity (dotted).", out.width="50%", fig.asp=1>>= 
@

>>>>

  To summarise, estimation of only $\bm{\alpha}_{\text{feat}}$ or
  $\bm{\alpha}_{\text{drug}}$ (and consequently $\phi_{jd}$ and $\chi_{d}$)
  by NIG$_{\text{f}}$ and NIG$_{\text{d}}$, respectively is relatively unbiased,
  as evident from simulation Scenarios 1 and 2 (Figures 
  \ref{fig:simulations_gdsc_est1} and \ref{fig:simulations_gdsc_est2}). 
  In contrast, simultaneous 
  estimation in Scenario 3 results in overestimated $\phi_{jd}$ and 
  underestimated $\chi_{d}$ by NIG$_{\text{f}+\text{d}}$
  (Figure \ref{fig:simulations_gdsc_est3}). We conjecture that this interplay
  of drug and feature variance components is due to near-unidentifiability.e
  r, the resultconsequences are limited, since theiances on the $\beta_{jd}$ level are unbiasleft ed 
  (Figure \ref{fig:simulations_gdsc_var}). If separately estimated, scale 
  parameters $\lambda_{\text{feat}}$ and $\lambda_{\text{drug}}$ are 
  overestimated by NIG$_{\text{f}}$ and NIG$_{\text{d}}$, respectively.
  If estimated simultaneously NIG$_{\text{f}+\text{d}}$ overestimates 
  $\lambda_{\text{feat}}$ and underestimated $\lambda_{\text{drug}}$. On the
  $\beta_{jd}$ level, $\bm{\lambda}$ influences kurtoses 
  $\mathcal{K}(\beta_{jd})$. Figure \ref{fig:simulations_gdsc_kurt} shows the
  mean estimated kurtoses versus their true values. Kurtoses seem to be
  underestimated by NIG$_{\text{f}+\text{d}}$. In contrast, 
  NIG$_{\text{f}+\text{d}}^{-1}$ overestimates the true mean of the
  $\mathcal{K}(\beta_{jd})$, \Sexpr{true.kurtb3}, at \Sexpr{kurtb3}.
  Lastly, the results from Scenario 3 in SM Section \ref{sm-sec:simulations} 
  show that the VB approximation is quite good as compared to standard MCMC.
<<simulations_gdsc_kurt, fig.cap="Simulation results for Scenario 3: mean estimated prior Kurtoses $\\mathcal{K}(\\beta_{jd})$ versus true values, with line of identity (dotted).", out.width="50%", fig.asp=1>>= 
@

### 2020-04-28 docs/supplement.Rnw
changed README file

tab3 <- sapply(1:ncol(tabm), function(s) {
  id <- which(tabm[, s]==min(tabm[, s], na.rm=TRUE))
  out <- paste0(round(tabm[, s], 3), " (", round(tabsd[, s], 3), ")")
  out[id] <- paste0("\\textbf{", out[id], "}")
  out})
colnames(tab3) <- c("$\\text{EMSE}_{\\text{total}}$", 
                    "$\\text{EMSE}_{\\text{top}}$", 
                    "$\\text{EMSE}_{\\text{bottom}}$", "$\\text{PMSE}$")     
rownames(tab3) <- c("NIG$_{\\text{f+d}}^-$", 
                    "NIG$_{\\text{f+d}}$", "ridge", "lasso")
@	
  The lasso and ridge models in this section were fit separately per drug, 
  using the \texttt{R} package \texttt{glmnet} 
  \cite[]{friedman_regularization_2010} with cross
	validated penalty parameters.

@@

  For completeness, Figure \ref{fig:simulations_gdsc_res5} displays samples
  from the full Bayesian NIG posterior, as introduced in Section 
  \ref{sec:gibbssampler}. Unfortunately, we were not able to generate enough
  samples for a reliable posterior using the Gibbs procedure as described in 
  the Section. We generated the samples from the full Bayesian posterior using
  stan through the R package \texttt{rstan} 
  \cite[]{guo_rstan:_2018} (see \url{https://github.com/magnusmunch/NIG} for 
  the implemented stan code).
<<simulations_gdsc_res5, fig.cap="Simulation results for Scenario 3: Samples from full Bayesian NIG posteriors (see Section \\ref{sec:gibbssampler}) with true value superimposed (red line) for (top row) group 1, (second row) group 2, (third row) group 3, and (bottom row) group 4 for both features and drugs, and (left column) $\\bm{\\phi}$ and (second column) $\\bm{\\chi}$.", out.width="100%">>= 
@

>>>>

tab3 <- sapply(1:ncol(tabm), function(s) {
  id <- which(tabm[, s]==min(tabm[, s], na.rm=TRUE))
  out <- paste0(round(tabm[, s], 3), " (", round(tabsd[, s], 3), ")")
  out[id] <- paste0("\\textbf{", out[id], "}")
  out})
colnames(tab3) <- c("$\\text{EMSE}_{\\text{total}}$", 
                    "$\\text{EMSE}_{\\text{top}}$", 
                    "$\\text{EMSE}_{\\text{bottom}}$", "$\\text{PMSE}$")     
rownames(tab3) <- c("NIG$_{\\text{f+d}}^-$", 
                    "NIG$_{\\text{f+d}}$", "ridge", "lasso")

n_eff.mean <- round(mean(summary.mcmc2$summary[, "n_eff"]), 0)
n_eff.min <- round(min(summary.mcmc2$summary[, "n_eff"]), 0)
time.mean <- round((time.mcmc2*1000/n_eff.mean)/3600, 0)
time.min <- round((time.mcmc2*1000/n_eff.min)/(7*24*3600), 0)
@	
  The lasso and ridge models in this section were fit separately per drug, 
  using the \texttt{R} package \texttt{glmnet} 
  \cite[]{friedman_regularization_2010} with cross
	validated penalty parameters.

@@

  For completeness, Figure \ref{fig:simulations_gdsc_res5} displays samples
  from the full Bayesian NIG posterior, as introduced in Section 
  \ref{sec:gibbssampler}. Unfortunately, we were not able to generate enough
  samples for a reliable posterior using the Gibbs procedure as described in 
  the Section. We generated the samples from the full Bayesian posterior using
  stan through the R package \texttt{rstan} 
  \cite[]{guo_rstan:_2018} (see \url{https://github.com/magnusmunch/NIG} for 
  the implemented stan code). The Figure shows that the posteriors for
  $\bm{\phi}$ and $\bm{\chi}$ have unexpected shapes and seem to have explored
  only a small part of the posterior space. This indication that
  posterior inference is not reliable is confirmed by the average and minimum
  effective sample sizes of \Sexpr{n_eff.mean} and \Sexpr{n_eff.min}over all 
  parameters, respectively. For reliable inference,
  \cite{kruschke_doing_2015} recommends an effective sample size of at least 
  10 000. Inspection of the computation times and a quick calculation learns 
  that we have to sample from the posterior
  for at least \Sexpr{time.mean} hours to obtain an average effective sample 
  size of 10 000. Moreover, to achieve a minimum effective sample size of 10 000 
  for all parameters, the sampler has to be run for about \Sexpr{time.min}
  weeks.
<<simulations_gdsc_res5, fig.cap="Simulation results for Scenario 3: Samples from full Bayesian NIG posteriors (see Section \\ref{sec:gibbssampler}) with true value superimposed (red line) for (top row) group 1, (second row) group 2, (third row) group 3, and (bottom row) group 4 for both features and drugs, and (left column) $\\bm{\\phi}$ and (second column) $\\bm{\\chi}$.", out.width="100%">>= 
@
